---
title: "Populated Places & Historical Markers in Middlesex and Suffolk, MA"
author: "@jahredwithanh"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE,  results='hide'}
library(leaflet)
library(sf)
library(maps)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(htmlwidgets)

us_places <- tidytuesdayR::tt_load("2023-06-27")

us_place_names <- us_places$us_place_names

us_place_history <- us_places$us_place_history

hist <- tidytuesdayR::tt_load('2023-07-04')

hist <- hist$historical_markers

knitr::opts_chunk$set(echo = TRUE)
```

The interactive map below shows U.S. populated places and historical markers in two counties in Massachusetts (Middlesex and Suffolk). Data are from the National Map Staged Products Directory (US Board of Geographic Names) and the [Historical Marker Database USA Index](https://www.hmdb.org/). County boundaries are also shown.

To change layers, click on the panel in the bottom right. Click a point to read historical information or access other links, if available.

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}

us_place_names <- us_place_names %>% filter(!is.na(prim_lat_dec) & !is.na(prim_long_dec))

us_place_names <- st_as_sf(us_place_names, coords = c("prim_long_dec", "prim_lat_dec"),
                           crs = "NAD83")

# suffolk places
suffolk_places <- us_place_names %>% 
  filter(state_name == "Massachusetts" & county_name == "Suffolk")

suffolk_places <- left_join(suffolk_places, us_place_history)

limits <- st_bbox(suffolk_places)

suffolk_places <- suffolk_places %>%
  mutate(map_desc = case_when(is.na(history) ~ paste("<B>", feature_name, "</B>", sep = ""),
                              !is.na(history) ~ paste("<B>", feature_name, "</B>:<br/>", history, sep = ""))
         )

# now Middlesex places:
middlesex_places <- us_place_names %>% 
  filter(state_name == "Massachusetts" & county_name == "Middlesex")

middlesex_places <- left_join(middlesex_places, us_place_history)

middlesex_places <- middlesex_places %>%
  mutate(map_desc = case_when(is.na(history) ~ paste("<B>", feature_name, "</B>", sep = ""),
                              !is.na(history) ~ paste("<B>", feature_name, "</B>:<br/>", history, sep = ""))
         )

# also quickly get county boundaries, then drop any outside boundaries
counties <- tidycensus::get_acs(geography = "county", variables = "B19013_001",
                                state = "MA", county = c("Middlesex", "Suffolk"),
                                geometry = TRUE)


suffolk_places <- suffolk_places[counties, ]
middlesex_places <- middlesex_places[counties, ]

# now clean the hist markers data
suffolk_hist <- hist %>% filter(grepl("^Suffolk", county_or_parish) & state_or_prov == "Massachusetts")

suffolk_hist <- st_as_sf(suffolk_hist, coords = c("longitude_minus_w", "latitude_minus_s"),
                           crs = "NAD83")
  
middlesex_hist <- hist %>% filter(grepl("^Middlesex", county_or_parish) & state_or_prov == "Massachusetts")

middlesex_hist <- st_as_sf(middlesex_hist, coords = c("longitude_minus_w", "latitude_minus_s"),
                           crs = "NAD83")

# format icons
Suffolk_hist_icons <- awesomeIcons(library = "ion",
                                   icon = "flag", markerColor = "green",
                                   iconColor = "black")
Middlesex_hist_icons <- awesomeIcons(library = "ion",
                                     icon = "flag", markerColor = "blue",
                                     iconColor = "black")

```

```{r, echo=FALSE, warning=FALSE, out.width = '100%'}
map1 <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  
  # counties
  addPolygons(data = counties, weight = 3,
              fillColor = "white", stroke = TRUE, 
              color = "black", group = "Counties",
              popup = counties$NAME)  %>%
  
  # points
  addCircleMarkers(data = suffolk_places, radius = 4, group = "Suffolk places",
                   weight = 3, color = "black", fillColor = "green", fillOpacity = 1,
                   label = suffolk_places$feature_name,
                   popup = suffolk_places$map_desc) %>%
  
  addCircleMarkers(data = middlesex_places, radius = 4, group = "Middlesex places",
                   weight = 3, color = "black", fillColor = "blue", fillOpacity = 1,
                   label = middlesex_places$feature_name,
                   popup = middlesex_places$map_desc) %>%
  
  addAwesomeMarkers(data = suffolk_hist, group = "Suffolk historical markers",
                    label = suffolk_hist$title,
                    popup = paste(suffolk_hist$location, "<br/><a href='", suffolk_hist$link, 
                                  "'>Learn more<a/>", sep = ""),
                    icon = Suffolk_hist_icons) %>%
  
  addAwesomeMarkers(data = middlesex_hist, group = "Middlesex historical markers",
                    label = middlesex_hist$title,
                    popup = paste(middlesex_hist$location, "<br/><a href='", middlesex_hist$link, 
                                  "'>Learn more<a/>", sep = ""),
                    icon = Middlesex_hist_icons) %>%
  
  # all layer control
  addLayersControl(
    position = "bottomright",
    overlayGroups = c("Suffolk places",  "Suffolk historical markers", 
                      "Middlesex places", "Middlsesex historical markers"),
    options = layersControlOptions(collapsed = TRUE)) %>%
   setMaxBounds(lng1 = as.numeric(limits[1]) - 0.75,
                lat1 = as.numeric(limits[2]) - 0.75,
                lng2 = as.numeric(limits[3]) + 0.75,
                lat2 = as.numeric(limits[4]) + 0.75
                )

map1
```

**#TidyTuesday week 26 + 27 \| [`@jahredwithanh`](https://twitter.com/jahredwithanh) \|**
